### 2023-09-30 latex/paper/main.tex
paper ray clusters

\pagebreak
\subsection{The Ray library}
The Ray library is an open-source distributed computing framework primarily designed for building scalable and high-performance applications. It was developed by the company Anyscale, which aimed to simplify the development of distributed and parallel computing applications. When it comes to matrix operation computation, using Ray for distributed computation offers several advantages over performing the operations on a single machine:
    \begin{itemize}
        \item \textbf{Faster execution}: With Ray you can use multiple machines and CPUs to compute operations. This significantly reduces the computation time compared to a single machine. Each machine can work on a subset of the matrix data, completing the calculations in parallel. This distributed approach can lead to substantial speedups.
        \item \textbf{Scalability}: As the size of the matrices grows, a single machine may struggle to handle the computational demands due to memory limitations or processing power constraints. Ray allows you to scale horizontally by adding more machines to the distributed setup.
        \item \textbf{Fault tolerance}: Ray offers fault tolerance mechanisms that ensure the continuity of computation even in the presence of failures. If a machine participating in the distributed computation fails, Ray can automatically redistribute the workload to other available machines.
        \item \textbf{Resource utilization}: With Ray, each machine contributes its processing power and memory capacity to the overall computation. This efficient utilization of resources allows you to make the most of the available hardware infrastructure, compared to a single machine that may be underutilized.
    \end{itemize}

@@

\subsection{Ray clusters}

\pagebreak
\subsection{The Matrix class}
Before implementing the parallelized algorithms for Ray, we write the code for the serial execution in the \textbf{Matrix} class (most of the functions defined in the class are auxiliary therefore were excluded for practical purposes):
\begin{lstlisting}[language=Python, label=Matrix class]
class Matrix:
    def __init__(self, data):
        if isinstance(data, list):
            self.data = [[round(val, 8) for val in row]
                                        for row in data]
        else:
            raise ValueError("Input must be a list")

>>>>

\subsection{The Ray library}
The Ray library is an open-source distributed computing framework primarily designed for building scalable and high-performance applications. It was developed by the company Anyscale, which aimed to simplify the development of distributed and parallel computing applications. When it comes to matrix operation computation, using Ray for distributed computation offers several advantages over performing the operations on a single machine:
\begin{itemize}
    \item \textbf{Faster execution}: With Ray you can use multiple machines and CPUs to compute operations. This significantly reduces the computation time compared to a single machine. Each machine can work on a subset of the matrix data, completing the calculations in parallel. This distributed approach can lead to substantial speedups.
    \item \textbf{Scalability}: As the size of the matrices grows, a single machine may struggle to handle the computational demands due to memory limitations or processing power constraints. Ray allows you to scale horizontally by adding more machines to the distributed setup.
    \item \textbf{Fault tolerance}: Ray offers fault tolerance mechanisms that ensure the continuity of computation even in the presence of failures. If a machine participating in the distributed computation fails, Ray can automatically redistribute the workload to other available machines.
    \item \textbf{Resource utilization}: With Ray, each machine contributes its processing power and memory capacity to the overall computation. This efficient utilization of resources allows you to make the most of the available hardware infrastructure, compared to a single machine that may be underutilized.
\end{itemize}

@@

\subsubsection{Ray clusters}
To run Ray applications on multiple nodes you must first deploy a \textbf{Ray cluster}. A Ray cluster is a set of worker nodes connected to a common Ray head node. These nodes can be physical machines, virtual machines, or containers running on a cloud infrastructure or a cluster manager like Kubernetes. Ray clusters can be fixed-size, or they may autoscale up and down according to the resources requested by applications running on the cluster.
\\\\
Each Ray cluster typically has one special node called the \textbf{head node}. This is where the main driver program is executed, which manages the cluster, schedules tasks, and communicates with the other nodes. The head node also hosts the Ray Dashboard, a web-based interface for monitoring and managing the cluster.
\\\\
The other nodes in the cluster are known as \textbf{workers}. These nodes are responsible for executing tasks and running jobs. Workers can be distributed across multiple machines and are managed by the head node. They execute user-defined Python functions and can be scaled up or down dynamically based on demand.
\\\\
A ray cluster is set up using a \textit{yaml} file, where all the details about the cluster configuration are defined. Once the file is ready, we simply launch \textit{ray up cluster.yaml} to start a ray instance on all worker nodes over ssh.
\begin{lstlisting}[caption={cluster.yaml}]
    cluster_name: distmat

    provider:
      type: local
      head_ip: "192.168.128.129"
      worker_ips:
        [
          "192.168.128.210",
          "192.168.128.211",
          "192.168.128.212",
          "192.168.128.213",
          "192.168.128.214",
          "192.168.128.220",
          "192.168.128.221",
          "192.168.128.222",
          "192.168.128.223",
          "192.168.128.224",
        ]

    auth:
      ssh_user: alo

    upscaling_speed: 1.0
    idle_timeout_minutes: 5
    head_start_ray_commands:
      - ray stop
      - ulimit -c unlimited && ray start --head --port=6379 --autoscaling-config=~/ray_bootstrap_config.yaml
    worker_start_ray_commands:
      - ray stop
      - ray start --address=$RAY_HEAD_IP:6379

\end{lstlisting}

\pagebreak
\subsection{The Matrix class}
Before implementing the parallelized algorithms for Ray, we write the code for the serial execution in the \textbf{Matrix} class (most of the functions defined in the class are auxiliary therefore were excluded for practical purposes):

\begin{lstlisting}[language=Python, label=Matrix class]
class Matrix:
    def __init__(self, data):
        if isinstance(data, list):
            self.data = [[round(val, 8) for val in row]
                                        for row in data]
        else:
            raise ValueError("Input must be a list")
